---
title: "R Notebook"
output: html_notebook
---
NOTE: interaction/hardcoding required for 2 variables:

1) population_list, this variable should be set == to the values in auto_pop_list, comment the pops you want to omit
2) pub_table_order, this variable should == the formated names held in "pop_IDs", order it how you wish the pops to be plotted

#Dependencies
```{r}
library(tictoc)
library("devtools", lib.loc="~/R/win-library/3.5")
library("genepop", lib.loc="~/R/win-library/3.5")
library("diveRsity", lib.loc="~/R/win-library/3.5")
library("adegenet", lib.loc="~/R/win-library/3.5")
library("xlsx", lib.loc="~/R/win-library/3.5")
library("sendplot", lib.loc="~/R/win-library/3.5")
library("dplyr", lib.loc="~/R/win-library/3.5")
library(reshape2)
library(ggplot2)
library(egg)
```
#genepop metadata
```{r}
#read genepop that we derived the admixture data from
#this will allow us to generalize our metadata in the figure
#e.g. n samples per population
SNPs_gen = read.genepop('../../v6_novoseq2_snps_genomic.gen', ncode = 3)
beep()


auto_pop_list = NULL
for (i in 1:n_distinct(SNPs_gen$pop)) {
  auto_pop_list = paste(auto_pop_list, "'",unique(SNPs_gen$pop)[i],"'",",", sep = "")
}
auto_pop_list
```

```{r}
#Paste your auto_pop_list strings here, they should be formatted as csv, then parse to different lines and comment out the ones you want to omit
population_list = c(
  'Big_Arbor_Vitae-10-13698',
  'Chippewa_Flowage-17-08347',
  # 'Cutfoot_Sioux-17-15586',
  'Delavan-15-01449',
  'Eau_Claire_River-17-01449',
  'Escanaba_Lake-17-02775',
  'Kawaguesaga-11-00300',
  # 'Lake_Koronis-17-15556',
  'Lake_Millicent-07-00697',
  'Lake_Wisconsin-15-05298',
  'Manitowish_Lake-17-10943',
  'Medicine_Lake-17-13247',
  # 'Mille_Lacs-17-15428',
  # 'Ottertail_Lake-17-15364',
  # 'Pike_River-17-15524',
  # 'Pine_River-17-15460',
  # 'Red_Lake-17-15396',
  'Sanford_Lake-17-02975',
  # 'Sarah_Lake-17-15332',
  # 'St_Louis_River-17-15492',
  'Turtle_Flambeau_Flowage-17-08397',
  'Willow_Flowage-11-00167',
  'WolfR_17-03648'
  )
snps_subset = SNPs_gen[which(SNPs_gen$pop %in% population_list)]

#extract n samples per population
n_by_pop = as.data.frame(matrix(NA,nrow = n_distinct(snps_subset$pop), ncol = 1))
for (i in 1:n_distinct(snps_subset$pop)) {
  n_by_pop[i,1] = length(which(snps_subset$pop == unique(snps_subset$pop)[i]))
}
#expand the point data to a format that will fit our genind
n_by_pop$V1 = as.double(n_by_pop$V1)
n_by_pop_strata = NULL
for (i in 1:nrow(n_by_pop)) {
  n_by_pop_strata = c(n_by_pop_strata,rep(x = n_by_pop[i,1],times = n_by_pop[i,1]))
}
n_by_pop_strata = as.data.frame(n_by_pop_strata)

#extract population names, removing the ID number that gets pulled in from the first sample
pop_IDs = snps_subset$pop
pop_IDs = gsub("\\d+","",pop_IDs)
pop_IDs = gsub("-","",pop_IDs)
pop_IDs = as.data.frame(pop_IDs)

#now merge the metadata with the genind
strat_dat = cbind.data.frame(n_by_pop_strata,pop_IDs)
strata(snps_subset) = strat_dat

#SNPs_gen$strata$n_by_pop_strata
#SNPs_gen@strata@pop_IDs

#initialize a palette of colors we can use downstream, 23 unique hexidecimal colors here
PCA23_palette = c("#000ffF","#0033CC","#006699","#009966","#006600",
"#66CC00","#CCFF00","#CC9900","#CC3300","#CC0033",
"#CC0099","#CC00FF","#6633FF","#CC66FF","#FF66CC",
"#FF3366","#FF6633","#FF9900","#6699ff","#33ccff","#00ffff","#00ff99")
```

#exploring ggplot option

this section requires a custom pallette for each level of K
the order of which K is which color moves from level to level so,
they have to be hard coded to facilitate the same cluster being represented with the same color across runs
```{r}
pdf("./SEMINAR_admixture_sorted.pdf", width = 11.5, height = 8, family = "Times")
#RAN 2-15
for (i in c(2,9,11)) {

read_name = paste(c("./admixture_all_pops/v6_novoseq2_snps_genomic.",as.character(i),".Q"),collapse = "")
tbl = read.table(read_name)

#requires ggplot2 and reshape pkgs
#transpose the results to have individuals in columns and K proportions in rows
tmp = as.data.frame(t(tbl))
#melt the data into column format, i.e., indv, K-proportion
tmp2 = melt(tmp)
rownames(tmp2) = c(1:nrow(tmp2))
#add name for each K value within an individual
tmp2$K_val = c(1:length(which(tmp2$variable == unique(tmp2$variable)[1])))

#add the pop ID data to each individual's K values
pop_IDs_for_admix = rep(pop_IDs$pop_IDs, each = i)

tmp2$pop = pop_IDs_for_admix

#making a list to sort the results on, this will organize the admixture plot by geography and not alphabetic
pub_table_order = c("Delavan","WolfR_","Lake_Wisconsin",
                    "Medicine_Lake","Willow_Flowage",
                    "Kawaguesaga","Big_Arbor_Vitae","Escanaba_Lake",
                    "Sanford_Lake","Manitowish_Lake","Turtle_Flambeau_Flowage",
                    "Chippewa_Flowage","Eau_Claire_River","Lake_Millicent",
                    "St_Louis_River",
                    "Sarah_Lake",
                    "Lake_Koronis",
                    "Mille_Lacs",
                    "Pine_River",
                    "Cutfoot_Sioux",
                    "Ottertail_Lake",
                    "Red_Lake",
                    "Pike_River"
                    )
#Loop through to subset populations in order of publish table. We're really just sorting here
tmp3 = NULL
for (j in 1:length(pub_table_order)) {
  tmp3 = rbind.data.frame(tmp3, tmp2[which(tmp2$pop == pub_table_order[j]),])
}
#add/change some metadata to make sure the plot is built in the order I wanted
rownames(tmp3) = c(1:nrow(tmp3))
tmp3$variable = rep(c(1:n_distinct(tmp2$variable)), each = i)
#rebuilding the x values for labels and lines
x_lab_vec = tmp3 %>% 
  group_by(pop) %>% 
  summarise(X_max = max(variable))
x_lab_vec = x_lab_vec %>% arrange(X_max)


############################
# This section builds on the sorting performed above, 
# in addtion to sorting based on the pop ID order (to match the order of tables)
# I'm sorting individuals within each pop based on the descending proportion of ancestry allocated to the major K for that pop
K_max = tmp3 %>% group_by(pop, K_val) %>% summarise(max_K = sum(value))
K_max2 = K_max %>% group_by(pop) %>% filter(max_K == max(max_K)) 
# need to update the variable ID as that is the Xval in the plot
# var_ID = as.integer(1)
admx_to_plot = NULL
for (p in pub_table_order) {
  # pull out a pop in the order of table
  test_sort = tmp3[which(tmp3$pop == p),]
  # extract the K value that accounts for the majority of that pop's ancestry
  K_iter = K_max2[which(K_max2$pop == p),"K_val"]
  # extract each individuals ancestry proporiton at that K
  test_sort2 = test_sort[which(test_sort$K_val == K_iter$K_val),]  
  # arrange the individuals based on their ancestry at that K
  test_sort3 = test_sort2 %>% arrange(desc(value,variable))
  # bind them back into the dataframe to plot  
  for (v in test_sort3$variable) {
    admx_to_plot = rbind.data.frame(admx_to_plot,tmp3[which(tmp3$variable == v),])
    # admx_to_plot[which(admx_to_plot$pop == p & admx_to_plot$variable == v),"variable"] = var_ID
    # var_ID = var_ID + 1
  }
}
xval = 1
var_ID_seq = seq(1,nrow(tmp3),by = i)
for (l in var_ID_seq) {
  admx_to_plot[c(l:(l+i-1)),"variable"] = xval
  xval = xval + 1
}

admx_to_plot



if (i == 9) {
  #                  Sanford,  Medicine, WISC/CHIP, CHIP,    Red     , Pike    , Delavan , Minnesota, WISC
PCA23_palette = c("#990099","#0099FF","#6600CC","#CC00CC","#FFCC00","#00FF00","#F79646","#00B050","#0066cc")
#custom colors
plt9 = ggplot() + geom_bar(aes(y = value, x = variable), fill = rep(PCA23_palette[1:i], nrow(pop_IDs)), data = admx_to_plot, stat = "identity")+
  geom_text(angle = 90, aes(x = x_lab_vec$X_max-15, y = 0.5, label = pub_table_order[which(pub_table_order %in% unique(tmp2$pop))]))+
  geom_vline(xintercept = x_lab_vec$X_max, color = "black", size=.5)+
  scale_fill_manual("legend", values = rep(PCA23_palette[1:i], nrow(pop_IDs)))+
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        plot.margin=unit(c(0,0,0,0), "cm"),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(color = "black"),
        axis.text.x = element_text(""),
        axis.title.x= element_blank(),
        axis.ticks.x= element_blank()
        # legend.position = "right"
        )+
  xlab("Individual")+
  ylab("Ancestry proportion")
}
if (i == 11) {
#                  Mille lacs, Medicine, Delavan, St Louis, WISC,   MINNESOTA, Sanford,   RED    , CHIP    ,  PIKE    , wisc/chip
PCA23_palette = c("#666600","#0099FF","#F79646","#9999FF","#0066cc","#00B050","#990099","#FFCC00","#CC00CC","#00FF00","#6600CC")
#custom colors
plt11 = ggplot() + geom_bar(aes(y = value, x = variable), fill = rep(PCA23_palette[1:i], nrow(pop_IDs)), data = admx_to_plot, stat = "identity")+
  geom_text(angle = 90, aes(x = x_lab_vec$X_max-15, y = 0.5, label = pub_table_order[which(pub_table_order %in% unique(tmp2$pop))]))+
  geom_vline(xintercept = x_lab_vec$X_max, color = "black", size=.5)+
  scale_fill_manual("legend", values = rep(PCA23_palette[1:i], nrow(pop_IDs)))+
      theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        plot.margin=unit(c(0,0,0,0), "cm"),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(color = "black"),
        axis.text.x = element_text(""),
        axis.title.x= element_blank(),
        axis.ticks.x= element_blank()
        # legend.position = "right"
        )+
  xlab("Individual")+
  ylab("Ancestry proportion")
}
if (i == 2) {
  PCA23_palette = c("#666600","#0099FF")
plt2 = ggplot() + geom_bar(aes(y = value, x = variable), fill = rep(PCA23_palette[1:i], nrow(pop_IDs)), data = admx_to_plot, stat = "identity")+
  geom_text(angle = 90, aes(x = x_lab_vec$X_max-15, y = 0.5, label = pub_table_order[which(pub_table_order %in% unique(tmp2$pop))]))+
  geom_vline(xintercept = x_lab_vec$X_max, color = "black", size=.5)+
  scale_fill_manual("legend", values = rep(PCA23_palette[1:i], nrow(pop_IDs)))+
      theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        plot.margin=unit(c(0,0,0,0), "cm"),
        axis.line = element_line(colour = "black"),
        axis.text = element_text(color = "black"),
        axis.text.x = element_text(""),
        axis.title.x= element_blank(),
        axis.ticks.x= element_blank()
        # legend.position = "right"
        )+
  xlab("Individual")+
  ylab("Ancestry proportion")
  ggarrange(plt2)
}
}
ggarrange(plt9,plt11, nrow = 2, ncol = 1)
dev.off()
```
